Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-12-07 paper_comparison/paper_comparison/paper_comparison.Rmd
clean the preprocessing code

```{r}
library(tidyverse)
# remotes::install_github("TengMCing/visage")
library(visage)

set.seed(10086)
```

```{r get-lineup-data}
if (!file.exists(here::here("data/vi_lineup.rds"))) {
  vi_lineup <- get_vi_lineup()
  saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))
} else {
  vi_lineup <- readRDS(here::here("data/vi_lineup.rds"))
}
```

```{r poly-conventional-simulation}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

if (!file.exists(here::here("data/poly_conventional_simulation.rds"))) {
  poly_conv_sim <- list()
  for (i in 1:100)
  {
    poly_conv_sim[[i]] <- map(1:2000,
                            function(x) {
                     shape <- sample(1:4, 1)
                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
                     x <- switch(x_dist,
                                 uniform = rand_uniform(-1, 1),
                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
                     mod <- poly_model(shape, x = x, sigma = e_sigma)
                     n <- sample(c(50, 100, 300), 1)
                     tmp_dat <- mod$gen(n)

                     es <- filter(vi_survey, 
                                  shape == {{shape}},
                                  e_sigma == {{e_sigma}},
                                  x_dist == {{x_dist}},
                                  n == {{n}}) %>%
                       head(1) %>%
                       pull(effect_size)

                     tibble(shape = shape,
                            e_sigma = e_sigma,
                            x_dist = x_dist,
                            n = n,
                            effect_size = es,
                            F_p_value = mod$test(tmp_dat)$p_value,
                            RESET3_p_value = mod$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
                            RESET4_p_value = mod$test(tmp_dat, test = "RESET", power = 2:4, power_type = "fitted")$p_value,
                            RESET5_p_value = mod$test(tmp_dat, test = "RESET", power = 2:5, power_type = "fitted")$p_value,
                            RESET6_p_value = mod$test(tmp_dat, test = "RESET", power = 2:6, power_type = "fitted")$p_value,
                            RESET7_p_value = mod$test(tmp_dat, test = "RESET", power = 2:7, power_type = "fitted")$p_value,
                            RESET8_p_value = mod$test(tmp_dat, test = "RESET", power = 2:8, power_type = "fitted")$p_value,
                            RESET9_p_value = mod$test(tmp_dat, test = "RESET", power = 2:9, power_type = "fitted")$p_value,
                            RESET10_p_value = mod$test(tmp_dat, test = "RESET", power = 2:10, power_type = "fitted")$p_value,
                            BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
                            SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
                            boot_id = i)
                   }) %>%
  reduce(bind_rows)
  }

@@

```{r heter-conventional-simulation}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

if (!file.exists(here::here("data/heter_conventional_simulation.rds"))) {
  heter_conv_sim <- list()
  for (i in 1:100)
  {
    heter_conv_sim[[i]] <- map(1:2000,
                            function(x) {
                     a <- sample(c(-1, 0, 1), 1)
                     b <- sample(c(0.25, 1, 4, 16, 64), 1)
                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
                     x <- switch(x_dist,
                                 uniform = rand_uniform(-1, 1),
                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
                                 even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
                     mod <- heter_model(a = a, b = b, x = x)
                     n <- sample(c(50, 100, 300), 1)
                     tmp_dat <- mod$gen(n)

                     es <- filter(vi_survey, 
                                  a == {{a}}, 
                                  b == {{b}},
                                  x_dist == {{x_dist}},
                                  n == {{n}}) %>%
                              head(1) %>%
                              pull(effect_size)

                     tibble(a = a,
                            b = b,
                            x_dist = x_dist,
                            n = n,
                            effect_size = es,
                            F_p_value = POLY_MODEL$test(mutate(tmp_dat, z = poly_model()$gen(n, computed = select(tmp_dat, x))$z))$p_value,
                            RESET3_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
                            RESET4_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:4, power_type = "fitted")$p_value,
                            RESET5_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:5, power_type = "fitted")$p_value,
                            RESET6_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:6, power_type = "fitted")$p_value,
                            RESET7_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:7, power_type = "fitted")$p_value,
                            RESET8_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:8, power_type = "fitted")$p_value,
                            RESET9_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:9, power_type = "fitted")$p_value,
                            RESET10_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:10, power_type = "fitted")$p_value,
                            BP_p_value = mod$test(tmp_dat)$p_value,
                            SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
                            boot_id = i)
                   }) %>%
  reduce(bind_rows)
  }

@@

Plotting the residuals against predicted values and each of the explanatory variables on a scatter plot is a recommend way to scrutinize their relationships. If there is any visually discoverable patterns, the model is potentially misspecified. However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans. It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none. [ET: I don't know if the former statement is true?] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

@@

The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers [ET: CITATION HERE?]. Section \ref{experimental-design} describes the experimental setup to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

## Visual test

Note that both the p-value and power comparison are not directly based on the diagnostic plots recommended by abundance of literature, but

@@

```{r plotsel, fig.cap=" A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50% of subjects evenly select other plots."}
vi_survey_with_plot %>%
  mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>%
  group_by(type) %>%
  mutate(n_eval = n()/20) %>%
  ungroup() %>%
  group_by(type, plot_p_value_rank) %>%
  summarise(weighted_sel = sum(weighted_sel), n_eval = first(n_eval)) %>%
  ggplot() +
  geom_col(aes(plot_p_value_rank, n_eval), fill = "grey", alpha = 0.3) +
  geom_col(aes(plot_p_value_rank, weighted_sel)) +
  facet_wrap(~type) +
  theme_light() +
  ylab("Weighted number of selections") +
  xlab("Plot p-value rank")
```

>>>>

```{r}
# OOP supports needed by `visuage`
# remotes::install_github("TengMCing/bandicoot")
# 
# Visual inference models and p-value calculation
# remotes::install_github("TengMCing/visage")

library(tidyverse)
library(visage)

# To control the simulation in this file
set.seed(10086)
```

```{r get-lineup-data}
# The lineup data used to draw residual plot needed to be downloaded
# from the github repo. Cache it.
if (!file.exists(here::here("data/vi_lineup.rds"))) {
  vi_lineup <- get_vi_lineup()
  saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))
} else {
  vi_lineup <- readRDS(here::here("data/vi_lineup.rds"))
}
```

```{r poly-conventional-simulation}
# Ensure the support of the regressor is [-1, 1]
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

# Run simulations to get the behaviours of conventional tests
if (!file.exists(here::here("data/poly_conventional_simulation.rds"))) {
  poly_conv_sim <- list()

  # 100 bootstrap samples
  for (i in 1:100)
  {
    poly_conv_sim[[i]] <- 
      # Every sample contains 2000 lineups
      map(1:2000, function(i) {

        # Sample a set of parameters
        shape <- sample(1:4, 1)
        e_sigma <- sample(c(0.5, 1, 2, 4), 1)
        x_dist <- sample(c("uniform", 
                           "normal", 
                           "lognormal", 
                           "even_discrete"), 1)
        x <- switch(x_dist,
                    uniform = rand_uniform(-1, 1),
                    normal = {
                      raw_x <- rand_normal(sigma = 0.3)
                      closed_form(~stand_dist(raw_x))
                      },
                    lognormal = {
                      raw_x <- rand_lognormal(sigma = 0.6)
                      closed_form(~stand_dist(raw_x/3 - 1))
                      },
                    even_discrete = rand_uniform_d(k = 5, even = TRUE))

        # Build the model
        mod <- poly_model(shape, x = x, sigma = e_sigma)

        # Sample the number of observations
        n <- sample(c(50, 100, 300), 1)

        # Generate data from the model 
        tmp_dat <- mod$gen(n)

        # Get the effect size from the pre-calculated table 
        es <- filter(vi_survey, 
                     shape == {{shape}},
                     e_sigma == {{e_sigma}},
                     x_dist == {{x_dist}},
                     n == {{n}}) %>%
          head(1) %>%
          pull(effect_size)

        # Return a data frame containing p-values of
        # F, RESET, BP and SW tests
        tibble(shape = shape,
               e_sigma = e_sigma,
               x_dist = x_dist,
               n = n,
               effect_size = es,
               F_p_value = mod$test(tmp_dat)$p_value,
               RESET3_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:3, 
                                         power_type = "fitted")$p_value,
               RESET4_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:4, 
                                         power_type = "fitted")$p_value,
               RESET5_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:5, 
                                         power_type = "fitted")$p_value,
               RESET6_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:6, 
                                         power_type = "fitted")$p_value,
               RESET7_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:7, 
                                         power_type = "fitted")$p_value,
               RESET8_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:8, 
                                         power_type = "fitted")$p_value,
               RESET9_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:9, 
                                         power_type = "fitted")$p_value,
               RESET10_p_value = mod$test(tmp_dat, 
                                          test = "RESET", 
                                          power = 2:10, 
                                          power_type = "fitted")$p_value,
               BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
               boot_id = i)
        }) %>%
  reduce(bind_rows)
  }

@@

```{r heter-conventional-simulation}
# Run simulations to get the behaviours of conventional tests
if (!file.exists(here::here("data/heter_conventional_simulation.rds"))) {
  heter_conv_sim <- list()

  # 100 bootstrap samples
  for (i in 1:100)
  {
    heter_conv_sim[[i]] <-
      # Every sample contains 2000 lineups
      map(1:2000, function(x) {

        # Sample a set of parameters
        a <- sample(c(-1, 0, 1), 1)
        b <- sample(c(0.25, 1, 4, 16, 64), 1)
        x_dist <- sample(c("uniform", 
                           "normal", 
                           "lognormal", 
                           "even_discrete"), 1)
        x <- switch(x_dist,
                    uniform = rand_uniform(-1, 1),
                    normal = {
                      raw_x <- rand_normal(sigma = 0.3)
                      closed_form(~stand_dist(raw_x))
                      },
                    lognormal = {
                      raw_x <- rand_lognormal(sigma = 0.6)
                      closed_form(~stand_dist(raw_x/3 - 1))
                      },
                    even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))

        # Build the model
        mod <- heter_model(a = a, b = b, x = x)

        # Sample the number of observations
        n <- sample(c(50, 100, 300), 1)

        # Generate data from the model 
        tmp_dat <- mod$gen(n)

        # Get the effect size from the pre-calculated table 
        es <- filter(vi_survey, 
                     a == {{a}}, 
                     b == {{b}},
                     x_dist == {{x_dist}},
                     n == {{n}}) %>%
          head(1) %>%
          pull(effect_size)

        # Return a data frame containing p-values of
        # F, RESET, BP and SW tests
        tibble(a = a,
               b = b,
               x_dist = x_dist,
               n = n,
               effect_size = es,
               F_p_value = POLY_MODEL$test(

                 # Create a pseudo z to be able to use F-test
                 tmp_dat %>%
                   mutate(z = poly_model() %>%
                            {.$gen(n, computed = select(tmp_dat, x))} %>%
                            pull(z))
                 )$p_value,
               RESET3_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:3, 
                                                power_type = "fitted")$p_value,
               RESET4_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:4, 
                                                power_type = "fitted")$p_value,
               RESET5_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:5, 
                                                power_type = "fitted")$p_value,
               RESET6_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:6, 
                                                power_type = "fitted")$p_value,
               RESET7_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:7, 
                                                power_type = "fitted")$p_value,
               RESET8_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:8, 
                                                power_type = "fitted")$p_value,
               RESET9_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:9, 
                                                power_type = "fitted")$p_value,
               RESET10_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:10, power_type = "fitted")$p_value,
               BP_p_value = mod$test(tmp_dat)$p_value,
               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
               boot_id = i)
        }) %>%
  reduce(bind_rows)
  }

@@

Plotting the residuals against predicted values and each of the explanatory variables on a scatter plot is a recommend way to scrutinize their relationships. If there is any visually discoverable patterns, the model is potentially misspecified. However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans.

It is especially common, particularly among new statistics students to report patterns when an experienced data analyst might quickly conclude that there are none. [ET: I don't know if the former statement is true? >> PL: change to students] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

@@

The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the types of departures that one expects to detect, and describes a formal statistical process for reading residual plots, called visual inference [ET: CITATION HERE? >> PL: arguments removed]. Section \ref{experimental-design} describes the experiment design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

<!-- ## Visual test  -->

<!-- Note that both the p-value and power comparison are not directly based on the diagnostic plots recommended by abundance of literature, but  -->

@@

```{r plotsel, fig.cap = "A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots."}
vi_survey_with_plot %>%
  mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>%
  group_by(type) %>%
  mutate(n_eval = n()/20) %>%
  ungroup() %>%
  group_by(type, plot_p_value_rank) %>%
  summarise(weighted_sel = sum(weighted_sel), n_eval = first(n_eval)) %>%
  ggplot() +
  geom_col(aes(plot_p_value_rank, n_eval), fill = "grey", alpha = 0.3) +
  geom_col(aes(plot_p_value_rank, weighted_sel)) +
  facet_wrap(~type) +
  theme_light() +
  ylab("Weighted number of selections") +
  xlab("Plot p-value rank")
```

### 2022-12-07 paper_comparison/paper_comparison/paper_comparison.tex
clean the preprocessing code

Plotting the residuals against predicted values and each of the
explanatory variables on a scatter plot is a recommend way to scrutinize
their relationships. If there is any visually discoverable patterns, the
model is potentially misspecified. However, correctly judging a residual
plot where no pattern exists is a painstakingly difficult task for
humans. It is especially common, particularly among new data analysts to
report patterns when an experienced data analyst might quickly conclude
that there are none. {[}ET: I don't know if the former statement is
true?{]} Generally, one looks for noticeable departures from the model
like non-linear dependency or heteroskedasticity. It is also possible to
conduct hypothesis tests for non-linear dependence
\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
\citep{breusch_simple_1979} for heteroskedasticity.

@@

The persistence of this advice to check the plots is curious, and
investigating why this might be common advice is the subject of this
paper. The paper is structured as follows. The next background section
describes the types of departures that one expects to detect, and
describes a formal process for reading residual plots, called visual
inference, that can avoid the concerns about subjectiveness of human
readers {[}ET: CITATION HERE?{]}. Section \ref{experimental-design}
describes the experimental setup to compare the decision made by formal
hypothesis testing, and how humans would read diagnostic plots. The
results are reported in Section \ref{results}. We finish with a
discussion on future work, in particular how the responsibility for
residual plot reading might be passed on to computer vision.

@@

\%\textbackslash end\{landscape\}

\hypertarget{visual-test}{%
\subsection{Visual test}\label{visual-test}}

Note that both the p-value and power comparison are not directly based
on the diagnostic plots recommended by abundance of literature, but

@@

>>>>

Plotting the residuals against predicted values and each of the
explanatory variables on a scatter plot is a recommend way to scrutinize
their relationships. If there is any visually discoverable patterns, the
model is potentially misspecified. However, correctly judging a residual
plot where no pattern exists is a painstakingly difficult task for
humans.

It is especially common, particularly among new statistics students to
report patterns when an experienced data analyst might quickly conclude
that there are none. {[}ET: I don't know if the former statement is
true? \textgreater\textgreater{} PL: change to students{]} Generally,
one looks for noticeable departures from the model like non-linear
dependency or heteroskedasticity. It is also possible to conduct
hypothesis tests for non-linear dependence \citep{ramsey_tests_1969},
and use a Breusch-Pagan test \citep{breusch_simple_1979} for
heteroskedasticity.

@@

The persistence of this advice to check the plots is curious, and
investigating why this might be common advice is the subject of this
paper. The paper is structured as follows. The next background section
describes the types of departures that one expects to detect, and
describes a formal statistical process for reading residual plots,
called visual inference {[}ET: CITATION HERE? \textgreater\textgreater{}
PL: arguments removed{]}. Section \ref{experimental-design} describes
the experiment design to compare the decision made by formal hypothesis
testing, and how humans would read diagnostic plots. The results are
reported in Section \ref{results}. We finish with a discussion on future
work, in particular how the responsibility for residual plot reading
might be passed on to computer vision.

@@

@@

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/plotsel-1}

\caption{A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots.}\label{fig:plotsel}
\end{figure}
